### 2024-09-29 project-plan.Rmd
la till metod

---
title: "Manuscript title here"
subtitle: "Add a Subtitle if Needed"
author: "Your name here"
output: 
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "out") })
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{=html}
<!-- The project plan should be 3-4 pages long and written in
Rmarkdown (like this document). Remove all comments in italic when you 
use this document as a template for your own project plan. -->
```
```{=html}
<!-- This is an R Markdown document. Markdown is a simple formatting
syntax for authoring HTML, PDF, and MS Word documents. For more
details on using R Markdown see http://rmarkdown.rstudio.com -->
```
```{=html}
<!-- This is a comment and will not be present in the compiled
document. Try "knitting" this document by clicking the knit button up
to the left, or by running `rmarkdown::render("project-plan.Rmd")` in
the console and you should only see headings -->
```
# Introduction

*This is part A) in the project plan [sv: projektredogörelse].*

*The background/introduction should summarise the scope of the problem,
what is known about the problem, what is not known, what specific
knowledge gap the study is supposed to fill and why this is important.
It should end be stating the aim of the study. The background should be
4-5 paragraphs long and each paragraph should be between 3-5 sentences.
Strive to make the paragraphs about the same length.*

*bibliography.bib includes an example reference. Add additional
references in that file as bibtex entries and then cite in text using:
"[@exampleKey9999]". One can convert DOIs' to bibtex entries using
<https://www.doi2bib.org/>. If you use Zotero you can simply export
entries to bibtex `Ctrl-Shift-C`, however you need to first setup Quick
Copy to export to bibtex
(<https://www.zotero.org/support/creating_bibliographies#quick_copy>)*

Trauma is a major global public health concern. It causes over four milion death and affects hundreds of millions each year, among those younger populations are the most affected [@noauthor_injuries_nodate]. The initial management of trauma patients once they reach hospital is very time sensitive and error prone [@mackersie_pitfalls_2010] [@teixeira_preventable_2007][@ivatury_patient_2008], and shock is the most common potentially preventable death within the first 24h of injury[@berry_shock_2015]. It is also the second major cause of early mortality in trauma patients [@kauvar_impact_2006], and patients in shock have a higher mortality rate than those not[@vang_shock_2022].

The definition for shock is circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are four types of causes, which are hypovolemic, obstructive, cardiogenic and distributive. These causes can occur alone or in combination. The most usual one during trauma are hypovolemic in form of a large bleeding[@teixeira_preventable_2007].

To differentiate the severity of trauma shock, we can divide shock into different degrees. One of such classification is the ATLS classification for major haemorrhage. The classification is divided into four classes, based on the patients: estimated bloodloss, heart frequency (HF), bloodpressure (BP), pulse pressure, respiratory rate (RR), mental status and urine output. However, this only meant to be used for hypovolemic shocks and not the other types. There is other classifications for predicting massive transfusion (TASH,TBSS, ABC), triaging (RETTS) or tracking the clinical deterioration of patients (NEWS2). The common denominator for these classifications is they are all roughly based on vital parameters, however none of them is specifically designed to classify trauma shock degree regardless of the type/cause of shock. ATLS in this regard, even though having its own shortcomings[@guly_vital_2011], still is one of the most used classifications for this topic(Källa).

One way to improve trauma care quality is trough identifying Opportunities for Improvement (OFI). This method has an advantage compared to traditional mortality reviews in that it also includes non-fatals. There are different processes to finding OFIs, but one of the primary ways is through structured multidisciplinary morbidity and mortality reviews of patient cases [@world_health_organization_guidelines_2009]. In order for the review content to be as comprehensive as possible for trauma care, the Donabedian quality of care framework can be used[@donabedian_quality_1988]. It is based on the three factors of structure, process and outcome. By following this framework, healthcare providers will be able to systematically find and address OFIs[källa?]. This will in turn improve the effectiveness of trauma care, directly impacting mortality and morbidity[källa?/8].

This study aims to describe the types of opportunities for improvement for adult trauma patients arriving in shock, and to assess how the degree of shock is associated with opportunities for improvement.

## Epidemiology

Trauma is a major global public health concern. It causes over four milion death and affects hundreds of millions each year, among those younger populations are the most affected (2). Globally low and middle income countries are disproportionately affected, but high income countries like Sweden also face considerable challenges(källa). In Sweden between 2018 and 2021, the most common age demographic for trauma is the working aged population, between 18 and 64 years old. (Swetrau ÅR, 2021). This not only leads to premature loss of life but also imposes a substantial socio-economic burden, due to long-term disabilities and rehabilitation needs.

## Trauma

Trauma are generally divided into two groups, blunt and penetrating injury. The most common type in Sweden is blunt injury, constituting around 90% of all trauma. The most common causes for blunt trauma are traffic related and falls, which account for roughly ½ and ⅓ of all blunt trauma(Swetrau ÅR, 2021).

Blunt and penetrating trauma have their own distinct clinical implications. Blunt trauma often results from impacts of different energy factor, and depending on the energy can result in anything from abbrations to fractures and internal organ damage. The most anatomically common type of blunt trauma is blunt abdominal trauma (BAT), at 75% of all blunt trauma(advances of abdominal trauma). The most common cause is traffic related such as, rapid deceleration leading to driver crashing into car furniture and causing contusions or rupture of internal organs.

Penetrating trauma on the other hand are often caused by gunshots or stab wounds(Emergency care and transportation of the sick and injured). The most common location is thorax and injuries there can lead to damage of the lungs and heart, which in turn interfere with breathing and circulation, such as pneumo/hemothorax and heart tamponade.

A consequence of both types of trauma is possibility of significant bleeding, which in turn can result in hemorrhagic shock.

## Shock

The definition for shock is circulatory failure leading to insufficient perfusion and oxygenation of tissue and organ(9). Some of the common symptoms are tachycardia, hypotenstion and altered mental status. Shock are usually classified into four different types based on physiological mechanism. The types are, hypovolemic, obstructive, cardiogenic and distributive, and they all have their own special characteristics(källa). These causes can occur alone or in combination, as there can be multiple disruptions on the circulatory system. Common examples of diagnoses for each type are, hypovolemic - bleeding, obstructive - pulmonary embolism, cardiogenic - STEMI  and, distributive - sepsis(källa).

The most common type of shock during trauma is the hypovolemic type due to a major hemorrhage(4). Hemorrhage leads to low blood volume, which beside the common central symptoms noted before, also results in peripheral vasoconstriction, which corresponds with cold, wet and pale extremities(källa).

## Shock classifications

To differentiate the severity of trauma shock, we can divide it into different degrees. One of such classification is the ATLS classification for major haemorrhage. ATLS stratifies patients into four classes, based on the patients: estimated blood loss, heart frequency (HF), bloodpressure (BP), pulse pressure, respiratory rate (RR), mental status and urine output. However, this only meant to be used for hypovolemic shocks and not the other types. There is other classifications for predicting massive transfusion (TASH,TBSS, ABC), triaging (RETTS) or tracking the clinical deterioration of patients (NEWS2). The common denominator for these classifications is they are all roughly based on vital parameters, however none of them is specifically designed to classify trauma shock degree regardless of the type/cause of shock. ATLS in this regard, even though having its own shortcomings(10), still is one of the most used classifications for this topic(Källa).

## Trauma quality improvement and oppertunities for improvement

The initial management of trauma patients once they reach hospital is very time sensitive and error prone (3) [(4)](5), and shock is the most common potentially preventable death within the first 24 hour of injury(6), as the midian time to death is 2 hours(Tisherman, 2015). To address these challenges and reduce preventable errors, different types of Quality Improvement (QI) initiatives were initiated. QI focuses on systematically evaluating healthcare processes and patient outcomes, aiming to reduce morbidity and mortality(WHO). By implementing formal quality measures and conducting collaborative case reviews, QI programs have strengthened trauma systems and improved patient outcomes in various settings globally (Juillard + Hashmi).

One way to improve trauma care quality is through identifying Opportunities for Improvement (OFI). This method has an advantage compared to traditional mortality reviews in that it also includes non-fatal outcomes, which gives a broader perspective and a chance to reduce chronic morbidity. There are different processes to finding OFIs, but one of the primary ways is through structured multidisciplinary morbidity and mortality (M&M) reviews of patient cases (11). In order for the review content to be as comprehensive as possible for trauma care, the Donabedian quality of care framework can be used(12). It is based on the three factors of structure, process and outcome. By following this framework, healthcare providers will be able to systematically find and address OFIs[källa?]. This will in turn improve the effectiveness of trauma care, directly impacting mortality and morbidity[källa?/8].

## Aims
This study aims to describe the types of opportunities for improvement for adult trauma patients arriving in shock, and to assess how the degree of shock is associated with opportunities for improvement.

# Methods

## Study design

We conducted a registry based retrospective cohort study, using data from the trauma registry and trauma care quality database at the Karolinska University Hospital in Solna.

## Setting

The trauma registry includes patients treated at Karolinska University Hospital in Solna between 2014 and 2023. The trauma care quality database is a subset of the trauma registry and includes patients selected for review. The Karolinska University Hospital in Solna treats all major trauma in the greater metropolitan area of Stockholm.

## Participants

We included all patients in trauma registry and Trauma care quality database. We excluded patients younger than 15 and/or were dead on arrival.

The inclusion criteria for the trauma registry is a patient either admitted through trauma team activation or a already admitted patient with an Injury Severity Score (ISS) greater than nine. Every trauma patient (registered in the trauma registry) at Karolinska University Hospital is then included in a morbidity and mortality review process, which involves both individual case evaluations by specialized nurses and audit filters. Patients identified with a high potential for OFIs are discussed at multidisciplinary conferences. The identified OFIs are then categorized into broader areas. The multidisciplinary conferences are held every six to eight weeks, during which an average of ten patient cases are reviewed by experienced specialists from all trauma-related fields. The presence or absence of OFIs is determined by consensus among all participants and is documented in the trauma care quality database.

## Variables and data sources/measurements

The outcome was defined as the presence of at least one OFI, determined by the multidisciplinary M&M conference in the “problem” registry. An OFI can be various types of errors, which are categorized into: clinical judgement error, inadequate resources, delay in treatment, missed injury, inadequate protocols, preventable death and other errors.

To classify the patients the following measurement will be used as independent variables for classification of shock regardless of their cause. Heart rate(HR), systolic blood pressure(SBP), respiratory rate(RR) and mental status. Based on these four measurements, the patients will be classified into four groups roughly based on the same value as in the ATLS trauma shock classifications.

Due to lack of data on FAST, type of injury and clinical chemistry, we will not be able to use categorisations such as TASH, TBSS and ABC with better sensitivity and specificity than ATLS[(13)](14)(15).
Due to the tenth edition of ATLS classifications lack actual numbers, we will be approximating numbers to the values. For HR and RR we will be using the numbers from the ninth edition of the classification. The SBP will be divded into class I and II where the SBP is >110, class III 110-90mmhg and class IV <90mmhg. This selection is based on tenth edition, where SBP is defined as normal in class I to II and normal/below in class III and clearly below normal in class IV. These two studies, redefined hypotension as 110mmhg respective 90mmhg dependent on age[(17)](18), therefore we choose class III as 110-90 mmhg and class IV below 90mmhg. We decided to not include Glasgow Coma Scale (GCS) due to difficulties in differentiating if any lowering is due to shock or traumatic brain injury (TBI). Urine output and pulse pressure were also neglected, due to lack of data.

To classify, all of the parameters in one class needs to be fulfilled. In the cases where the parameters belong in different classes, we will be classifying according to the most severe parameter in the same way as Dunham et al(16). At same time ATLS manual states one of the common pitfalls is “Diagnosis of shock can be missed when only a single parameter is used.” therefore to improve sensitivity we decided to use the worst parameter for classification.

We will be only using the base deficit parameter according to the tenth edition of the ATLS classification for haemorrhage. There are three reasons. The first one is ATLS tenth edition lack numerical values for vital parameters. The second one is that bace deficit is superior to vital parameters as a predictive parameter for mortality(16). The third one is due to nature of having multiple parameters, makes it difficult to categorise a patient with vital parameters in two or more classes and difficult to separate lower GCS due to traumatic brain injury from shock as cause. The classes will be divided into class I: 0 - (-2), class II: (-2)-(-6), class III: (-6)-(-10), and class IV: <(-10).

## Bias

The trauma care quality database is build up in a way where every error can be classified into a certain type of OFI, however for paitent that died there is only the category “preventable death”. Therefore, survivorship bias may occur as we dont have the detail which type of OFI actually occured.

## Study size

All available data in trauma registry and trauma care quality database will be included.

## Quantitative variables

## Statistical methods

The statistical analysis will be performed using R, a programming language and environment for statistical computing. We will present the type of OFI as percentage distributions and then visualize it in a pie chart. Unadjusted and adjusted logistic regression will be used to determine the association between the OFI, degree/class of shock according to ATLS and other patient factors such as age, sex, preinjury ASA and ISS. It will be presented as odds ratios (OR) between presence of OFI, and shock classes. The OR will be determined with 95% confidence intervals, and a significance level of 5% will be used. All statistical analysis will first be done on synthetic data and later implemented on the data collected from the trauma registry and the trauma care quality database to ensure objectivity.

# Ethics

The handling of registry data and the extraction of information from medical records pose a potential risk to patient privacy. To mitigate this, patients were anonymized after data collection, with personal identity numbers replaced by unique study identifiers.

From a self-determination standpoint, the Swedish trauma registry does not require patient consent for data registration. However, it is mandatory for the caregiver to inform patients of their inclusion in the registry. Patients have the right to view, correct, opt out of, and request the deletion of their data, without any impact on the care they receive.

In this study, data will be presented at the group level, ensuring that no individual patient details can be identified. Ethical approvals has been obtained for this study, with the record numbers: 2021-02541, 2021-03531 and 2023-02975-02.

The benefits of this study, in relation to its risks, are substantial due to its potential to improve care for future critically ill patients. This research is a step toward reducing mortality and morbidity in a significant portion of the younger and economically active population worldwide.

# Project Update and Timeline

*This is part B) in the project plan [sv: projektredogörelse].*

*Save the image generated from the Gantt chard spreadsheet into the
assets folder, rename it to project-gantt.png*

![**The Gantt chart will be displayed here when you place it in the
assets folder under "project-gantt.png"**](assets/project-gantt.png)

Doing right now: - Introduction - Methods

To do: - Coding to sort data - Coding for regression - Analysis - Results - Discussion -
Conclusion and Abstract

# Back-up Plan

After discussion with my supervisor, it seems like the part which have the highest possibility of taking too much time is coding, and therefore he and other cosupervisor will assist with coding if needed.

# References

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->