### 2024-11-10 manuscript.Rmd
förde över metod från theisis och tog bort ethics

---
title: "Manuscript title here"
author: "Your name here"
output:
  html_document: default
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{=html}
<!-- This is an R Markdown document. Markdown is a simple formatting
syntax for authoring HTML, PDF, and MS Word documents. For more
details on using R Markdown see http://rmarkdown.rstudio.com -->
```
```{=html}
<!-- This is a comment and will not be present in the compiled
document. Try "knitting" this document by clicking the knit button up
to the left, or by running `rmarkdown::render("manuscript.Rmd")` in
the console and you should only see headings -->
```
```{=html}
<!-- Structure and most comments are from the Strengthening the
reporting of observational studies in epidemiology (STROBE) statement,
see https://www.strobe-statement.org/. For more explanations see the
paper Strengthening the Reporting of Observational Studies in
Epidemiology (STROBE):explanation and elaboration.  -->
```
```{=html}
<!-- Indicate the study’s design with a commonly used term in the
title, for example: "Associations between resuscitation interventions
and opportunities for improvement in adult trauma patients: A cohort
study" -->
```
# Abstract

```{=html}
<!-- Provide in the abstract an informative and balanced summary of
what was done and what was found. Not more than 300 words. -->
```
## Background

## Methods

## Results

## Conclusion

# Introduction

```{=html}
<!-- Explain the scientific background and rationale for the
investigation being reported. End by stating the aim or specific
objectives, including any prespecified hypotheses. The introduction
should not be longer than 1-1.5 pages, typically 4-5 paragraphs, each
6-8 lines long. Support every statement with a citation -->
```
Trauma is a major global public health concern. It causes over four milion death and affects hundreds of millions each year, among those younger populations are the most affected [@noauthor_injuries_nodate]. The initial management of trauma patients once they reach hospital is very time sensitive and error prone [@mackersie_pitfalls_2010] [@teixeira_preventable_2007][@ivatury_patient_2008], and shock is the most common potentially preventable death within the first 24h of injury[@berry_shock_2015]. It is also the second major cause of early mortality in trauma patients [@kauvar_impact_2006], and patients in shock have a higher mortality rate than those not[@vang_shock_2022].

The definition for shock is circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are four types of causes, which are hypovolemic, obstructive, cardiogenic and distributive. These causes can occur alone or in combination. The most usual one during trauma are hypovolemic in form of a large bleeding[@teixeira_preventable_2007].

To differentiate the severity of trauma shock, we can divide shock into different degrees. One of such classification is the ATLS classification for major haemorrhage. The classification is divided into four classes, based on the patients: estimated bloodloss, heart frequency (HF), bloodpressure (BP), pulse pressure, respiratory rate (RR), mental status and urine output. However, this only meant to be used for hypovolemic shocks and not the other types. There is other classifications for predicting massive transfusion (TASH,TBSS, ABC), triaging (RETTS) or tracking the clinical deterioration of patients (NEWS2). The common denominator for these classifications is they are all roughly based on vital parameters, however none of them is specifically designed to classify trauma shock degree regardless of the type/cause of shock. ATLS in this regard, even though having its own shortcomings[@guly_vital_2011], still is one of the most used classifications for this topic(Källa).

| **Parameter**                     | **Class I** | **Class II** (Mild) | **Class III** (Moderate) | **Class IV** (Severe)        |
|-----------------------------------|-------------|----------------------|---------------------------|-------------------------------|
| Approximate blood loss            | <15%        | 15–30%              | 31–40%                    | >40%                          |
| Heart rate                        | ↔           | ↔/↑                 | ↑                         | ↑/↑↑                          |
| Blood pressure                    | ↔           | ↔                   | ↔/↓                       | ↓                             |
| Pulse pressure                    | ↔           | ↓                   | ↓                         | ↓                             |
| Respiratory rate                  | ↔           | ↔                   | ↔/↑                       | ↑                             |
| Urine output                      | ↔           | ↔                   | ↓                         | ↓↓                            |
| Glasgow Coma Scale score          | ↔           | ↔                   | ↓                         | ↓                             |
| Base deficit                      | 0 to –2 mEq/L | –2 to –6 mEq/L     | –6 to –10 mEq/L          | –10 mEq/L or less             |
| Need for blood products           | Monitor     | Possible            | Yes                       | Massive Transfusion Protocol  |

One way to improve trauma care quality is trough identifying Opportunities for Improvement (OFI). This method has an advantage compared to traditional mortality reviews in that it also includes non-fatals. There are different processes to finding OFIs, but one of the primary ways is through structured multidisciplinary morbidity and mortality reviews of patient cases [@world_health_organization_guidelines_2009]. In order for the review content to be as comprehensive as possible for trauma care, the Donabedian quality of care framework can be used[@donabedian_quality_1988]. It is based on the three factors of structure, process and outcome. By following this framework, healthcare providers will be able to systematically find and address OFIs[källa?]. This will in turn improve the effectiveness of trauma care, directly impacting mortality and morbidity[källa?/8].

This study aims to describe the types of opportunities for improvement for adult trauma patients arriving in shock, and to assess how the degree of shock is associated with opportunities for improvement.

# Methods

## Study design

We conducted a registry based retrospective cohort study, using data from the trauma registry and trauma care quality database at the Karolinska University Hospital in Solna.

## Setting

The trauma registry includes patients treated at Karolinska University Hospital in Solna, which treats all major trauma in the greater metropolitan area of Stockholm. We included patients registered between 2014 and 2023. The trauma care quality database is a subset of the trauma registry and includes patients selected for review.

## Participants

The inclusion criteria for the trauma registry is a patient either admitted through trauma team activation or an already admitted patient with an Injury Severity Score (ISS) greater than nine. Every trauma patient (registered in the trauma registry) at Karolinska University Hospital is then included in a morbidity and mortality review process, which involves both individual case evaluations by specialized nurses and audit filters. Patients identified with a high potential for OFIs are discussed at multidisciplinary conferences. The identified OFIs are then categorized into broader areas. The presence or absence of OFIs is determined by consensus among all participants and is documented in the trauma care quality database.

We included all patients in the trauma registry and Trauma care quality database. We excluded patients younger than 15 and/or were dead on arrival.

## Variables and data sources/measurements

The outcome was defined as the presence of at least one OFI, determined by the multidisciplinary M&M conference in the trauma care quality registry. An OFI can be various types of errors, which are categorized into: clinical judgment error, inadequate resources, delay in treatment, missed injury, inadequate protocols, preventable death and other errors.
The patients will be classified to different degrees of shock in parallel and separately, once with systolic blood pressure(SBP) and once with base excess (BE), regardless of their cause of shock. Based on these two parameters, the patients will be classified into four classes roughly based on the ATLS trauma shock classifications.
We included five other metrics to be adjusted for, due to potential for confounding. Pre-injury ASA and gender were categorical, meanwhile age, ISS and INR were kept continuous as they were directly from the register.

## Bias

## Study size

All available data in trauma registry and trauma care quality database will be included.

## Quantitative variables

The BE parameter was divided into four classes with these cutoffs: Class I (above -2), Class II (-2 to -6), Class III (-6 to -10), and Class IV (below -10). Similarly, SBP was divided into four classes: Class I (above 110 mmHg), Class II (109-100 mmHg), Class III (99-90 mmHg), and Class IV (below 90 mmHg).
This classification follows the tenth edition ATLS hemorrhage classification, which does not specify exact values for any parameter other than BE. To solve this, we assigned numerical SBP values to each class based on findings from Eastridge et al. and Oyetunji et al., who redefined hypotension as 110 mmHg respective 90 mmHg dependent on age[@eastridge_hypotension_2007; @oyetunji_redefining_2011]. Therefore, we used 110 mmHg as the lower limit for (Class I: no shock) - (ATLS: normal SBP) and below 90 mmHg for (Class IV: clear shock) - (ATLS: clear hypotension), class 2 and 3 divided equally in between. This also aligns with the classification done by Mutschler et al.[@mutschler_critical_2013].

**ATLS tenth edition - Systolic blood pressure and Base excess (BD)**

| **Parameter - original ATLS 10th edition**| **Class I** | **Class II** (Mild)  | **Class III** (Moderate)  | **Class IV** (Severe) |
|-----------------------------------|-------------|----------------------|---------------------------|-------------------------------|
| Base deficit                      | 0 to –2 mEq/L | –2 to –6 mEq/L     | –6 to –10 mEq/L           | –10 mEq/L or less             |
| Systolic Blood pressure           | ↔           | ↔                    | ↔/↓                       | ↓                             |
| **Parameter - numerical approximated SBP**                                                                                         |
| Systolic Blood pressure           | >110 mmhg   | 109-100 mmhg         | 99-90 mmhg                | <90  mmhg                     |

Please note that ATLS classification defines class I for the BE parameter as (0 to -2) however, we choose to follow its original source[@mutschler_renaissance_2013], which defines class I as base deficit (Inverted BE) ≤ 2, with no lower limit (i.e. an upper limit for BE).

## Statistical methods

The statistical analysis will be performed using R, a programming language and environment for statistical computing. We will present the types of OFI as percentage distributions and then visualize them in a bar chart. Unadjusted logistic regression will be used to determine the association between the OFI, and degree of shock defined separately by classifying BE and SBP roughly according to ATLS classification of hemorrhagic shock. Adjusted logistic regression will then incorporate other patient factors such as age, sex, preinjury ASA, INR, and ISS. It will be presented as odds ratios (OR) between the presence of OFI, and shock classes. The OR will be determined with 95% confidence intervals, and a significance level of 5% will be used. All statistical analysis will first be done on synthetic data and later implemented on the data collected from the trauma registry and the trauma care quality database to ensure objectivity. Missing data will be addressed by listwise deletion.

# Results
```{r main, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
source("main.R") ## This "imports" the main script file of your project and run any code in it
```

```{=html}
<!--

Participants
------------

(a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

(b) Give reasons for non-participation at each stage

(c) Consider use of a flow diagram

There were a total of `r nrow(merged.data)` patients in the trauma registry. Out of these, `r nrow(study.sample)` patients had been reviewed for the presence of opportunities for improvement. After exluding patients younger than 15 and/or were dead on arrival, `r nrow(study.sample)` patients remained. A total of X patients were finally excluded due to missing data, leaving us with Y patients for the final analysis.

Descriptive data
----------------

(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

(b) Indicate number of participants with missing data for each
variable of interest

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)

Sample characteristics are presented in Table \@ref(tab:sample-characteristics). A total of `r ofi` patients had at least one opportunity for improvement. The median age was `r age`. Most patients were male `r gender`.

Outcome data
------------

Cohort study — Report numbers of outcome events or summary measures
over time

Case-control study — Report numbers in each exposure category, or
summary measures of exposure

Cross-sectional study — Report numbers of outcome events or summary
measures

Main results
------------

(a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period

Other analyses 
--------------

Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->
```
You can include code in this document like this:

```{r sample-characteristics, echo=FALSE}
sample.characteristics.table
```

You can also embed plots:

```{r plot, echo=FALSE}
plot(pressure)
```

You can also mix text and code, so called inline code, like this: `r 2+5`.

# Discussion

```{=html}
<!--

Key results
-----------
Summarise key results with reference to study objectives

Limitations
-----------

Discuss limitations of the study, taking into account sources of
potential bias or imprecision.  Discuss both direction and magnitude
of any potential bias

Interpretation
--------------

Give a cautious overall interpretation of results considering
objectives, limitations, multiplicity of analyses, results from
similar studies, and other relevant evidence

Generalisability
----------------

Discuss the generalisability (external validity) of the study results

-->
```
# Conclusion

<!-- Give a short conclusion, corresponding to the aim -->

# References

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->