### 2025-07-13 manuscript.Rmd
la in förändringarna enligt google docs och källorna i discussion

---
title: Opportunities for improvement in the care of adult trauma patients arriving
  in shock
author: "Carl Skog"
output:
  #word_document: default
  #bookdown::pdf_document2:
    #latex_engine: xelatex
  html_document: default
bibliography: bibliography.bib
csl: vancouver.csl
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{=html}
<!-- This is an R Markdown document. Markdown is a simple formatting
syntax for authoring HTML, PDF, and MS Word documents. For more
details on using R Markdown see http://rmarkdown.rstudio.com -->
```

```{=html}
<!-- This is a comment and will not be present in the compiled
document. Try "knitting" this document by clicking the knit button up
to the left, or by running `rmarkdown::render("manuscript.Rmd")` in
the console and you should only see headings -->
```

```{=html}
<!-- Structure and most comments are from the Strengthening the
reporting of observational studies in epidemiology (STROBE) statement,
see https://www.strobe-statement.org/. For more explanations see the
paper Strengthening the Reporting of Observational Studies in
Epidemiology (STROBE):explanation and elaboration.  -->
```

```{=html}
<!-- Indicate the study’s design with a commonly used term in the
title, for example: "Associations between resuscitation interventions
and opportunities for improvement in adult trauma patients: A cohort
study" -->
```

# Abstract

```{=html}
<!-- Provide in the abstract an informative and balanced summary of
what was done and what was found. Not more than 300 words. -->
```

Background: Hemorrhagic shock is the leading cause of preventable trauma
deaths. Opportunities for improvement (OFI) are common in deaths from
massive bleeding. Aims: This study aimed to describe the types of OFI
for adult trauma patients arriving in shock and assess the association
between shock severity and OFI. Methods: A retrospective cohort study
was conducted on adult trauma patients treated at Karolinska University
Hospital (2014-2023). Shock severity was categorized into four classes
using base excess (BE) and systolic blood pressure (SBP). OFI identified
during morbidity and mortality conference were analyzed using bi- and
multivariable logistic regression, with adjustments for age, sex,
American Society of Anesthesiologists Classification (ASA),
international normalized ratio (INR), and injury severity score (ISS)
Results: 4919 patients were included, of which 281 had OFI. Clinical
judgment error was the most common OFI type for patients in shock.
Unadjusted analyses showed positive associations between shock
severities and OFI. However, after adjusting for ISS, patients with
severe shock (BE \< -10) showed 63% reduced odds of OFI compared to
those without shock (p = 0.026). ISS was a significant confounder (p \<
0.001). No significant association was observed between SBP classes and
OFI in adjusted analysis. Conclusion: Association between shock and OFI
seems to be nonlinear, with moderate shock (BE (-2) – (-10)) having
highest odds of OFI. The lack of significance in SBP classification
highlights potential framework limitations compared to BE. These
findings emphasize the importance of accounting for ISS and using
validated shock classifications in future research.

## Background

## Methods

## Results

## Conclusion

# Introduction

```{=html}
<!-- Explain the scientific background and rationale for the
investigation being reported. End by stating the aim or specific
objectives, including any prespecified hypotheses. The introduction
should not be longer than 1-1.5 pages, typically 4-5 paragraphs, each
6-8 lines long. Support every statement with a citation -->
```

# Introduction

Trauma is a major global public health concern. It causes over four
million deaths, among those younger populations are the most
affected[@noauthor_injuries_nodate]. Beyond the high mortality rate,
trauma also imposes a substantial socio-economic burden, due to
long-term disabilities and rehabilitation
needs[@noauthor_injuries_nodate].

The initial management of trauma patients is highly time-sensitive and
prone to error [@mackersie_pitfalls_2010; @teixeira_preventable_2007;
@ivatury_patient_2008]. Quality Improvement (QI) initiatives address
these challenges by systematically evaluating care processes and
outcomes, aiming to reduce morbidity and mortality
[@world_health_organization_guidelines_2009].

Hemorrhagic shock is, the second most common cause of death following
trauma and it, is the leading preventable cause of death within the
first 24 hours post-injury [@kauvar_impact_2006; @berry_shock_2015;
@teixeira_preventable_2007], and opportunities for improvement are
common in patients who die because of massive traumatic bleeding
[@oreilly_opportunities_2013].

This may indicate that these preventable events potentially leading to
suboptimal outcomes are common also in patients who survive massive
bleeding. This study aims to describe the types of opportunities for
improvement for adult trauma patients arriving in shock, and to assess
how the degree of shock is associated with opportunities for
improvement. This knowledge would help clinicians avoid such events in
bleeding trauma patients.

# Methods

## Study design

We conducted a registry based retrospective cohort study, using data
from the trauma registry and trauma care quality database at the
Karolinska University Hospital in Solna.

## Setting

The trauma registry is comprised of patients treated at the Karolinska
University Hospital in Solna, which treats all major trauma in the
greater metropolitan area of Stockholm. We included patients registered
between 2014 and 2023. The trauma care quality database is a subset of
the trauma registry and includes patients selected for review.

## Participants

Inclusion in the trauma registry requires either admission through
trauma team activation or presenting with an Injury Severity Score (ISS)
greater than nine after admission to the Karolinska University Hospital.
Each trauma patient is then included in a morbidity and mortality review
process, which involves both individual case evaluations by specialized
nurses and audit filters. Patients identified with a high potential for
OFIs are discussed at multidisciplinary conferences. An OFI can be
various types of preventable events, and can be categorized as: clinical
judgment error, inadequate resources, delay in treatment, missed injury,
inadequate protocols, preventable death and other errors. The presence
or absence of OFIs is determined by consensus among all participants and
is documented in the trauma care quality database.

All patients in the trauma registry and trauma care quality database
were included. Patients younger than 15 and/or were dead on arrival were
excluded.

## Variables and data sources/measurements

The outcome was defined as the presence of at least one OFI, determined
by the multidisciplinary M&M conference in the trauma care quality
registry. An OFI can be various types of preventable events, and can be
categorized as: clinical judgment error, inadequate resources, delay in
treatment, missed injury, inadequate protocols, preventable death and
other errors.

The patients will be classified to different degrees of shock in
parallel and separately, once with systolic blood pressure (SBP) and
once with base excess (BE), regardless of their cause of shock. Based on
these two parameters, the patients will be classified into four classes
roughly based on the ATLS trauma shock classifications (See Table
\@ref(tab:atls) [@noauthor_atls_2018]).

| **Parameter** | **Class I** | **Class II** (Mild) | **Class III** (Moderate) | **Class IV** (Severe) |
|----------------|--------------|--------------|--------------|--------------|
| Approximate blood loss | \<15% | 15–30% | 31–40% | \>40% |
| Heart rate | ↔ | ↔/↑ | ↑ | ↑/↑↑ |
| Blood pressure | ↔ | ↔ | ↔/↓ | ↓ |
| Pulse pressure | ↔ | ↓ | ↓ | ↓ |
| Respiratory rate | ↔ | ↔ | ↔/↑ | ↑ |
| Urine output | ↔ | ↔ | ↓ | ↓↓ |
| Glasgow Coma Scale score | ↔ | ↔ | ↓ | ↓ |
| Base deficit | 0 to –2 mEq/L | –2 to –6 mEq/L | –6 to –10 mEq/L | –10 mEq/L or less |
| Need for blood products | Monitor | Possible | Yes | Massive Transfusion Protocol |

: (#tab:atls) ATLS classification for haemorrhagic shock, adapted from
[@noauthor_atls_2018].

We included five other metrics to be adjusted for, due to potential for
confounding. Pre-injury ASA and gender were categorical, meanwhile age,
ISS and INR were kept continuous.

## Study size

All available data in trauma registry and trauma care quality database
will be included.

## Quantitative variables

Table \@ref(tab:mod-atls) shows how we defined shock according to BE and
SBP. The BE parameter was divided into four classes with these cutoffs:
Class I (above -2), Class II (-2 to -6), Class III (-6 to -10), and
Class IV (below -10). Similarly, SBP was divided into four classes:
Class I (above 110 mmHg), Class II (109-100 mmHg), Class III (99-90
mmHg), and Class IV (below 90 mmHg). This classification follows the
tenth edition ATLS hemorrhage classification [@noauthor_atls_2018],
which does not specify exact values for any parameter other than BE. To
solve this, we assigned numerical SBP values to each class based on
findings from Eastridge et al. and Oyetunji et al., who redefined
hypotension as 110 mmHg respective 90 mmHg dependent on
age[@eastridge_hypotension_2007; @oyetunji_redefining_2011]. Therefore,
we used 110 mmHg as the lower limit for (Class I: no shock) - (ATLS:
normal SBP) and below 90 mmHg for (Class IV: clear shock) - (ATLS: clear
hypotension), class 2 and 3 divided equally in between. This also aligns
with the classification done by Mutschler et
al.[@mutschler_critical_2013].

| **Parameter - original ATLS 10th edition** | **Class I** | **Class II** (Mild) | **Class III** (Moderate) | **Class IV** (Severe) |
|----------------|--------------|--------------|--------------|--------------|
| Base deficit | 0 to –2 mEq/L | –2 to –6 mEq/L | –6 to –10 mEq/L | –10 mEq/L or less |
| Systolic Blood pressure | ↔ | ↔ | ↔/↓ | ↓ |
| **Parameter - numerical approximated SBP** |  |  |  |  |
| Systolic Blood pressure | \>110 mmhg | 109-100 mmhg | 99-90 mmhg | \<90 mmhg |

: (#tab:mod-atls) ATLS tenth edition - Systolic blood pressure and Base
excess (BD).

Please note that ATLS classification defines class I for the BE
parameter as (0 to -2) however, we choose to follow its original
source[@mutschler_renaissance_2013], which defines class I as base
deficit (Inverted BE) ≤ 2, with no lower limit (i.e. an upper limit for
BE).

## Statistical methods

The statistical analysis was performed using R, a programming language
and environment for statistical computing. OFI subcategories were
presented as percentage distributions and visualized in a bar chart.
Bivariable logistic regression was used to determine the unadjusted
association between the OFI, and degree of shock defined by BE and SBP
classifications. Multivariable logistic regression incorporated and
adjusted for other patient factors such as age, sex, preinjury ASA, INR,
and ISS. It was presented as odds ratios (OR) between the presence of
OFI, and shock classes. The OR was determined with 95% confidence
intervals, and a significance level of 5% was used. All statistical
analysis was first done on synthetic data and then later implemented on
the data collected from the trauma registry and the trauma care quality
database to ensure objectivity. Missing data was addressed by listwise
deletion.

# Results

```{r main, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
source("main.R") ## This "imports" the main script file of your project and run any code in it
```

## Participants

There was a total of 14022 patients in the trauma registry. After
excluding the patients younger than 15 and/or were dead on arrival,
there were 12153 patients remaining. Out of those, 7152 patients had
been reviewed for the presence of OFI. A total of 2233 patients were
excluded due to missing data, resulting in 4919 patients for the final
analysis. The selection process is shown in Figure 1. The variable with
the most missing data is BE, for which data was missing in 1721
patients. Detailed data regarding missing data is displayed in the
supplementary material.

Figure 1. Flowchart of selection of study population.

## Descriptive data

Descriptive data is shown in Table 3. The median age was 39 years (25,
56), and 3506 (71%) of the individuals were males. Clinically, the
median ISS was 5, which is considered minor injuries, and the median INR
was 1.00, and the most common Preinjury ASA class was 1, at 62%.

Overall, 3989 (81%) of the patients were classified as having no shock
(class 1) according to the BE classification, the sum of the remaining
classes accounted for 930 (19%) patients. The distribution of the
classes was decreasing in numbers with increased severity, with class 2:
644 (13%), class 3: 170 (3.5%), and class 4: 116 (2.4%).

According to SBP, 4455 (91%) patients had no shock (class 1), while the
remaining classes with patients in shock accounted for 464 (9%) of all
the patients. Among the classes defined with shock, class 2 was the
biggest at 239 (4.9%), followed by class 4 at 127 (2.6%), and lastly
class 3 with 99 (2.0%).

Table 3

```{r sample-characteristics-reg, echo=FALSE}
sample.characteristics.table_reg |>
  modify_caption("Sample characteristics used in regression.")
```

## Outcome data

Out of the 4919 patients reviewed without missing data, 281 (5.7%)
patients had at least one OFI. The most common OFI type was clinical
judgment error, numbering 103 (37%) patients. The other three major OFI
types were inadequate resources for 61 (22%) patients, delay in
treatment for 49 (17%) patients, and missed diagnosis for 48 (17%)
patients. The least common OFI was “other errors”, which was present in
5 (1.8%) patients.

The most common OFI in patients without chock (class 1) was clinical
judgment error both in the BE and SBP groups, while the other OFI types
also followed a similar distribution between the groups. Clinical
judgment error was the most common OFI in the BE group class 2, 3 and 4,
at 34.6%; 31.2%; and 42.9% respectively. On the other hand, the largest
OFI type in SBP class 2 was delay in treatment at 50%. The frequencies
of clinical judgement error and missed diagnosis were similar in the SBP
class 3 group, at 40% each. Lastly, the most common OFI in class 4 of
the SBP group was clinical judgment error, at 46.7%.

------------------------------------------------------------------------

Table 4 - Sub-OFI, classified according to BE

```{r sub-ofi BE, echo=FALSE}
BEsubofi_tbl
```

Table 5 - Sub-OFI, classified according to SBP

```{r sub-ofi SBP, echo=FALSE}
SBPsubofi_tbl
```

## Main results

The unadjusted analysis of the association between shock class categorised by BE is shown in Table 6. This analysis showed that compared to class 1 (no shock), the odds of OFI in class 2 and 3 were increased by 61% (OR 1.61, 95% CI 1.17-2.20, p-value 0.003) and 91% (OR 1.91, 95% CI 1.08-3.16, p-value 0.018) respectively. Table 7 shows the association between shock classes categorised by SBP. Compared to class 1 (no shock), the odds of OFI in class 4 were increased by 128% (OR 2.28, 95% CI 1.26-3.85, p-value 0.004).

The adjusted analysis was conducted in two steps. In the first step, we adjusted for age, gender, preinjury ASA score, and INR, but not for ISS. Compared to the unadjusted analysis, this initial adjustment led to minor changes in ORs. In the BE group, the OR for class 2 changed by approximately 1% (from 1.61 to 1.60), and for class 3, it changed by about 2% (from 1.91 to 1.93), as shown in Table 6. In the SBP group, the OR for class 4 changed by around 7% (from 2.28 to 2.21), as shown in Table 7.

In the second step, ISS was added to the adjustment model to complete the fully adjusted model. For the BE group (Table 6), this led to a shift in statistical significance: instead of class 2 and 3, class 4 became statistically significant. In this fully adjusted model, class 4 was associated with 63% lower odds of an opportunity for improvement (OR 0.37, 95% CI 0.14–0.83, p = 0.026). In contrast, for the SBP group (Table 7), no class remained statistically significant in the fully adjusted model.

Table 6 adjusted log BE

```{r masterBE, echo=FALSE}
master_combined_table_stepBE
```

Table 7 adjusted log SBP

```{r masterSBP, echo=FALSE}
master_combined_table_stepSBP
```

# Discussion

```{=html}
<!--

Key results
-----------
Summarise key results with reference to study objectives

Limitations
-----------

Discuss limitations of the study, taking into account sources of
potential bias or imprecision.  Discuss both direction and magnitude
of any potential bias

Interpretation
--------------

Give a cautious overall interpretation of results considering
objectives, limitations, multiplicity of analyses, results from
similar studies, and other relevant evidence

Generalisability
----------------

Discuss the generalisability (external validity) of the study results

-->
```

## Key results

We found that the most common type of OFI is clinical judgment error across almost all shock severities. The unadjusted models suggested a positive association between moderate shock and OFI; adjustment reversed this, with severe shock showing 63% reduced odds.

## Limitations

The study design is limited by the lack of data on heart rate, FAST, detailed types of injury, and clinical chemistry. This resulted in choosing a proprietary simplified ATLS framework instead of categorizations such as TASH, TBSS, and ABC with better sensitivity and specificity than the ATLS framework [@yucel_trauma_2006;@ogura_modified_2016;@hanlin_prehospital_2024]. Another limitation is that missing data were handled by listwise deletion, which reduced the sample size by 2233 and may also introduce bias due to missing BE values. Also worth noting is that the sample size for most of the shock classes (class 2 to 4) are roughly around 100 to 200 patients, whereas those with positive OFI are as low as five patients. This may have resulted in higher randomness and inflated effect sizes.

## Interpretation

The lower odds of OFI for patients with BE < -10 may be explained by two theories. One is that more severe injuries may receive heightened attention and more protocol-driven responses. This would in turn, result in reduced risk of deviation from optimal care (and thus OFIs). This theory is also partly supported by the observation that the OFIs “inadequate protocols” and “missed diagnosis” generally occur with lower frequencies with higher shock degrees, regardless of whether classed with BE or SBP. Do note that the low frequencies of OFI in the severe shock classes, limit this theory with higher chance of randomness.

The second theory is survival bias, where the reduced odds of OFI may be explained by that this study only incorporated the surviving patients. This may result in only the most optimally cared patients surviving. Patients with BE < -10 also have high mortality rates [@mutschler_renaissance_2013]. Hypothetically, this would result in the severe classes of shock, with more dying patients with OFIs being eliminated from the data, due to death as the outcome.

The reversal of OR and distribution of statistical significance may be attributed to multicollinearity, as both BE and ISS assess injury severity. Theoretically, anatomical injury, bleeding, and physiological response should be closely interconnected. A correlation between ISS and BD (inverted BE) or SBP was also shown in the two studies by Mutscher et al. [@mutschler_critical_2013;@mutschler_renaissance_2013]. Therefore, adjusting for each other may generate results that are difficult to interpret. This may also explain why classes 2 and 3 are no longer statistically significant, whereas class 4 remains significant.

One other factor which may have contributed to this pattern of results, especially the distribution of significant classes and not as much the reversal after adjustment with ISS, was that ISS was included as a continuous (linear) variable in the model. As shown by Albaaj et al. [@albaaj_patient_2023], increases in ISS do not correspond proportionally to increases in the likelihood of OFI. Therefore, its relationship with opportunities for improvement (OFI) is likely not linear, which results in different statistical significances between the classes.

### Comparison to literature

A previous study by O’Reilly et al [@oreilly_opportunities_2013], which looked into opportunities for performance improvements (OPI) in 7511 cases, whereof 127 patients whose death was affected by hemorrhage, found that the most common OPI was related to decision making (63 out of 150 OPIs). This aligns with our findings that patients in shock (Class 2-4) were most affected by the OFI, “clinical judgment error” (BE 34.7%; SBP 35.3%).

No studies were found that specifically examined OFIs in patients in hemorrhagic shock, but studies on OFIs in general trauma populations show results in a similar theme. A study by T.P. Saltzherr [@saltzherr_preventability_2011] from a Dutch trauma center reported that “delay in treatment” or “error in management” was the most common error type. Similarly, Martin et al. [@martin_analysis_2009] found that delays to surgery were significant contributors to adverse outcomes, for a military hospital in Baghdad. Sanddal et al [@sanddal_analysis_2011] identified the main issues of care for a trauma system in development in Utah as chest injury diagnostic and management, airway management, and fluid resuscitation. Teixeira et al. [@teixeira_preventable_2007], analyzing preventable deaths in a mature trauma center, also confirmed the impact of delayed treatment and poor clinical decision-making.

In comparison, our study found that “clinical judgment error” was the most common OFI for all patients and patients in shock. While the relative ranking may vary, whether delay in treatment or clinical judgment error is most common, these two categories consistently emerged as the most frequent across studies. This also suggests that the types of OFIs that were common in the general trauma population were similarly frequent in patients in shock. Moreover, this pattern appears consistent regardless of the studied population, whether the focus was on preventable deaths or on patients who survived their injuries.

# Conclusion

Clinical judgment error was the most common OFI in shock patients. The association between shock and OFI does not seem to be linear, but rather the moderate shock has higher odds of OFI compared to the mild and severe shocks. These findings emphasize the importance of improved protocols for initial management of non-critical patients to minimise decision error.

# Contributions
Data collection from the trauma registry, ethics application and study designed by Martin Gerdin Wärnberg. Carl sorted, programmed and analyzed the data, and then wrote the thesis with support from Martin Gerdin Wärnberg, Johanna Berg, Jonatan Attergrim and Kelvin Szolnoky.

# References

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->